Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems

### 2023-02-26 rmd/manuscript.Rmd
edit: manuscript
result, one paragraph

We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL, such as ecosystem size (number of patches), complexity (branching probability), and disturbance regime. In addition, our simulation model incorporates intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL across extrinsic factors. We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors. Our model is an important step in unifying drivers of FCL across extrinsic and intrinsic drivers and parsing their context-dependency in spatially complex ecosystems.

@@

## Overall pattern

There were three consistent patterns in the relationship between FCL and ecosystem properties. First, both ecosystem size and complexity had a stronger positive effect in high disturbance regime. Second, the effects of ecosystem complexity was stronger than ecosystem size in general. Third, omnivory had a negative influence on the FCL regardless of simulation scenarios. Below, we report our results for low and high disturbance regimes separately.

## **Low disturbance regime**

@@

>>>>

We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL, such as ecosystem geometry [size (number of patches) and complexity (branching probability] and disturbance regime. In addition, our simulation model incorporates intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL across extrinsic factors. We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors. Our model is an important step in unifying drivers of FCL across extrinsic and intrinsic drivers and parsing their context-dependency in spatially complex ecosystems.

@@

## Overall effects of ecosystem geometry

There were three consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity). First, positive effects of ecosystem size and complexity were generally found when both disturbance frequency and intensity were high. Second, omnivory plays a central role in modulating the effects of ecosystem geometry on the FCL. Third, resource availability and dispersal did not change the qualitative relationship between the FCL and ecosystem geometry, although they had important direct effects on the FCL.

## Ecosystem geometry interacts with disturbance

Ecosystem geometry interacted with disturbance to dictate FCL in branching ecosystems. With no disturbance, both geometric features (size and complexity) had limited influence on FCL. However, as disturbance frequency increases, the relationships between FCL and ecosystem geometry became positive. The exception is the "chain" scenario (i.e., no omnivory). When this food web structure was combined with no disturbance, ecosystem size had a clear positive effect on FCL while ecosystem complexity had a negative influence. These opposing influences of ecosystem size and complexity may reflect their contrasting impacts on the total carrying capacity within a network (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it.

@@

## Ecosystem geometry interacts with omnivory

## Productivity and dispersal

##
